---
title: "Data Cleaning and Summary"
author: "Havisha Khurana"
date: "11/2/2021"
output: 
    html_document:
        highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's load packages
```{r packages and theme, include = FALSE}
packages <- c("here", "janitor", "tidyverse", "skimr", "rio", "psych")
lapply(packages, library, character.only = TRUE)
theme_set(theme_minimal())
```

Let's read the data.
```{r read data}
hnb <- import(here("data","Harassment and Bullying.csv")) %>% 
    as_tibble()
hnb_col <- colnames(hnb)
```

Quick summary: 

* There are `r ncol(hnb)` variables in this dataset.   
* First 8 are school characteristics  
* 9-13 are Harassment and Bullying allegations -- Allegations based on sex, ethnicity, disability status, sexual orientation and religion. Total varianles: `r length(grep("SCH_HBALLEGATIONS", hnb_col))`  
* 14 - 79 are #students reported as being bullied by different intent and demographic characteristics. Total variables: `r length(grep("SCH_HBREPORTED", hnb_col))`  
* 80 - 145 are #students discilined for bullying or harassing by different intent and demographic characteristics. Total variables: `r length(grep("SCH_DISCIPLINED", hnb_col))`  
* Totals are available (starts with TOT) for different intent (rex, race and disability status), reported/disciplined and gender (male, female) combo. Total variables for reported: `r length(grep("TOT_HBREPORTED", hnb_col))`, and Total variables for disciplined: `r length(grep("TOT_HBDISCIPLINED", hnb_col))`   

_Note to self:_ `grep()` saves the indices of columns

First, let's change all negative numbers to NA. 

```{r hnb_pos}
remove_neg <- function (x) ifelse(x < 0, NA, x)
hnb_pos <- hnb %>% 
    mutate(across(c(9:145), remove_neg))
```

Nice!

Let's make a shorter dataset to work with. I have allegations, reported and disciplined for the following intents: sex, race and disability. Let's just take these forward.

```{r hnb_sub}
cat_sum <- function(x) rowSums(select(hnb_pos, starts_with(x)), na.rm = TRUE)

hnb_sub <- hnb_pos %>% 
    select("LEA_STATE","LEA_NAME","COMBOKEY", starts_with(c("SCH_HBALLEGATIONS","TOT_HBREPORTED","TOT_HBDISCIPLINED"))) %>% 
    mutate(TOT_HBREPORTED_SEX = cat_sum("TOT_HBREPORTED_SEX"),
           TOT_HBREPORTED_RAC = cat_sum("TOT_HBREPORTED_RAC"),
           TOT_HBREPORTED_DIS = cat_sum("TOT_HBREPORTED_DIS"),
           TOT_HBDISCIPLINED_SEX = cat_sum("TOT_HBDISCIPLINED_SEX"),
           TOT_HBDISCIPLINED_RAC = cat_sum("TOT_HBDISCIPLINED_RAC"),
           TOT_HBDISCIPLINED_DIS = cat_sum("TOT_HBDISCIPLINED_DIS"),
           TOT_HBALLEGATIONS = cat_sum("SCH_HBALLEGATIONS"),
           TOT_HBREPORTED = cat_sum("TOT_HBREPORTED"),
           TOT_HBDISCIPLINED = cat_sum("TOT_HBDISCIPLINED")) %>% 
    filter(if_any(c(27:29),~ . > 0))
```

That looks like a neat data set. We have total `r ncol(hnb_sub)` variables - 3 are identifiers and rest are numeric. Also, we have filtered out rows where all numeric columns are 0.

Let's find some summary of cols.

```{r summary}
skim(hnb_sub)

```

This is super interesting! Some conclusions: 

* Allegations are highest, followed by disciplined and then reported. Probably, some teacher/adult saw the incident and disciplined the students involved, and hence things were not reported. Or, one student reported multiple people, hence the whole group was disciplined.    
* Allegation, reported and disciplined have highest raw count for the 'Disability' intent, followed by 'sex' and 'race'. However, the mean in all categories follow the order: 'sex', 'race' and 'disability'
* In reported, often female students report more frequencty for 'sex' intent, but male students report more often for 'race' and 'disability'
* In disciplined, for all intents, male students are more often disciplined than female students.

Let's do some plots now.

```{r scatter plot matrix allegations, out.width="100%"}

pairs.panels(hnb_sub[,4:8], 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
             )

#pairs(hnb_sub[,4:8], pch = 19, lower.panel = NULL)
#plotmatrix(hnb_sub[,27:29], colour="gray20")
```

```{r total, out.width="100%"}
pairs.panels(hnb_sub[,27:29], 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
             )
```
        
```{r group total, out.width="100%"}
pairs.panels(hnb_sub[,21:26], 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
             )
```

Let's look at state-wide boxplot, but each point represents the district

```{r state and district, out.width="100%"}
hnb_sub2 <- hnb_sub %>% 
    group_by(LEA_NAME) %>% 
    mutate(district_all = mean(TOT_HBALLEGATIONS),
           district_rep = mean(TOT_HBREPORTED),
           district_dis = mean(TOT_HBDISCIPLINED))

hnb_sub2 %>% 
    group_by(LEA_STATE) %>% 
    ggplot() +
    geom_boxplot(aes(LEA_STATE, district_all))
```

