---
title: "CCRC EnrollmentCleaning"
author: "Manuel Vazquez"
date: "11/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading necesarry packages. Using pacman to help load a list of packages. The if statement
is required in case it needs to be installed in the computer of somebody running this.

```{r }
if (!require("pacman")) install.packages("pacman") 
pacman::p_load(tidyverse, rio, here, ggridges, readr, ggthemes, sandwich, pracma, lme4,equatiomatic, performance, jtools, sundry, purrr,naniar,stargazer,vtable, reshape2, skimr, rddtools,tidylog,janitor)
```
<center>
# Part 1 - Data Import and light Cleaning
</center>


## 1.a Enrollment File 

In this section, I will import the enrollment file and subset it to keep relevant variables. 
This file will be the base file. 

```{r }
#Importing the Enrollment file
enroll<- read_csv(here("data", "Enrollment.csv")) %>%
  as_tibble() %>%
  subset(select = -c(9:31))
```

Exploring the file - I notice that there are some minimum values that are less than
zero, negative values don't make sense for enrollment. Ned to replace those to 
missing.

```{r }
skim(enroll)
``` 


First, need to replace values that have negative values with a missing value, then
creating enrollment counts for race/ethnicity groups as well as EL and SPED.

```{r }
enroll <- enroll %>%
  mutate_all(~replace(., . < 0, NA)) %>%
  mutate(enroll_all = TOT_ENR_M + TOT_ENR_F,
         enroll_latinx = SCH_ENR_HI_M + SCH_ENR_HI_F,
         enroll_aian = SCH_ENR_AM_M + SCH_ENR_AM_F,
         enroll_asianam = SCH_ENR_AS_M + SCH_ENR_AS_F,
         enroll_nhpi = SCH_ENR_HP_M + SCH_ENR_HP_F,
         enroll_black = SCH_ENR_BL_M + SCH_ENR_BL_F,
         enroll_white = SCH_ENR_WH_M + SCH_ENR_WH_F,
         enroll_twomore = SCH_ENR_TR_M + SCH_ENR_TR_F,
         enroll_el = SCH_ENR_LEP_M + SCH_ENR_LEP_F,
         enroll_elserv = TOT_LEPPROGENR_M + TOT_LEPPROGENR_F,
         enroll_sped = SCH_ENR_IDEA_M + SCH_ENR_IDEA_F)
```

Checking to see if the code worked by checking the distribution of all the variables. 
In theory, none should go below zero in these box graphs. Based on the results, it appears
as if this worked out well 

```{r }
enroll %>% 
  melt(id.vars = "SCH_NAME", 
       measure.vars = c("enroll_all", "enroll_latinx", "enroll_aian", "enroll_asianam",
                        "enroll_nhpi", "enroll_black", "enroll_white", "enroll_twomore",
                        "enroll_el", "enroll_elserv", "enroll_sped"),
       variable.name = "Group",
       value.name = "Count") %>%
  ggplot(aes(x = Count, y = Group)) +
  geom_boxplot() + 
  theme_minimal()

```

Now creating a set of new variables. I am creating enrollment percents for each
race/ethnicity group as well as EL and SPED.

```{r }
enroll <- enroll %>%
  mutate(enroll_latinx_pct = (enroll_latinx/enroll_all)*100,
         enroll_aian_pct = (enroll_aian/enroll_all)*100,
         enroll_asianam_pct = (enroll_asianam/enroll_all)*100,
         enroll_nhpi_pct = (enroll_nhpi/enroll_all)*100,
         enroll_black_pct = (enroll_black/enroll_all)*100,
         enroll_white_pct = (enroll_white/enroll_all)*100,
         enroll_twomore_pct = (enroll_twomore/enroll_all)*100,
         enroll_el_pct = (enroll_el/enroll_all)*100,
         enroll_elserv_pct = (enroll_elserv/enroll_all)*100,
         enroll_sped_pct = (enroll_sped/enroll_all)*100)
```

Also checking to see distribution of all the variables by ploting them. I notice some 
are reporting percents higher than 100. For example, some schools, the percent of students
who are EL enrolled is greater than 100. This shouldn't be the case.

```{r }
enroll %>% 
  melt(id.vars = "SCH_NAME", 
       measure.vars = c( "enroll_latinx_pct", "enroll_aian_pct", "enroll_asianam_pct",
                        "enroll_nhpi_pct", "enroll_black_pct", "enroll_white_pct", "enroll_twomore_pct",
                        "enroll_el_pct", "enroll_elserv_pct", "enroll_sped_pct"),
       variable.name = "Group",
       value.name = "Count") %>%
  ggplot(aes(x = Count, y = Group)) +
  geom_boxplot() + 
  theme_minimal()

enroll %>%
    filter(enroll_el_pct > 100) %>%
    select(SCH_NAME,SCH_ENR_LEP_F,SCH_ENR_LEP_M, enroll_el, enroll_all, enroll_el_pct)
```

Recoding those schools that report percentages higher than 100. Doing this for EL% and SPED%.
For those cases, I just recoded back to 100 percent. It appears like it worked based on the visual.
Afterwards, I subset the file to only keep the school information, enrollment totals, and percent enrollment.

```{r }


enroll <- enroll %>%
  mutate(enroll_el_pct = ifelse(enroll_el_pct > 100,100,enroll_el_pct),
         enroll_elserv_pct = ifelse(enroll_elserv_pct > 100,100,enroll_elserv_pct),
         enroll_sped_pct = ifelse(enroll_sped_pct > 100,100,enroll_sped_pct)) 
```
  
Creating  District Level Independent variables (% EL, % SPED, % students of color)

```{r }
enroll <- enroll %>%
  group_by(LEAID) %>%
  mutate(dist_enrollment = sum(enroll_all),
         dist_percent_white = (sum(enroll_white)/dist_enrollment)*100,
         dist_percent_latinx = (sum(enroll_latinx)/dist_enrollment)*100,
         dist_percent_black = (sum(enroll_black)/dist_enrollment)*100,
         dist_percent_twomore = (sum(enroll_twomore)/dist_enrollment)*100,
         dist_percent_asianam= (sum(enroll_asianam)/dist_enrollment)*100,
         dist_percent_nhpi = (sum(enroll_nhpi)/dist_enrollment)*100,
         dist_percent_aian = (sum(enroll_aian)/dist_enrollment)*100,
         dist_percent_el = (sum(enroll_el)/dist_enrollment)*100,
         dist_percent_sped = (sum(enroll_sped)/dist_enrollment)*100
         )


enroll %>%
  melt(id.vars = "SCH_NAME", 
       measure.vars = c( "dist_percent_white", "dist_percent_latinx", "dist_percent_black",
                        "dist_percent_twomore", "dist_percent_asianam", "dist_percent_nhpi",
                        "dist_percent_aian", "dist_percent_el", "dist_percent_sped"),
       variable.name = "Group",
       value.name = "Percent") %>%
  ggplot(aes(x = Percent, y = Group)) +
  geom_boxplot() + 
  theme_minimal()
  


enroll %>%
  ggplot(aes(x = dist_enrollment, y = LEA_STATE_NAME)) +
  geom_density_ridges() + theme_minimal()
```

Replacing district rates with values greater than 100 to just 100

```{r }

enroll <- enroll %>%
  mutate(dist_percent_el = ifelse(dist_percent_el > 100,100,dist_percent_el),
         dist_percent_sped = ifelse(dist_percent_sped > 100,100,dist_percent_sped))

enroll %>%
  melt(id.vars = "SCH_NAME", 
       measure.vars = c( "dist_percent_white", "dist_percent_latinx", "dist_percent_black",
                        "dist_percent_twomore", "dist_percent_asianam", "dist_percent_nhpi",
                        "dist_percent_aian", "dist_percent_el", "dist_percent_sped"),
       variable.name = "Group",
       value.name = "Percent") %>%
  ggplot(aes(x = Percent, y = Group)) +
  geom_boxplot() + 
  theme_minimal()
``` 
  
Making the dataset smaller and just keeping variables that we need  
  
```{r }  
enroll <- enroll %>%  
select(c("LEA_STATE", "LEA_STATE_NAME", "LEAID", "LEA_NAME", "SCHID", "SCH_NAME", 
          "COMBOKEY", "JJ",  "enroll_all", "enroll_latinx", "enroll_aian", "enroll_asianam",
           "enroll_nhpi", "enroll_black", "enroll_white", "enroll_twomore",
           "enroll_el", "enroll_elserv", "enroll_sped","enroll_latinx_pct", 
          "enroll_aian_pct", "enroll_asianam_pct", "enroll_nhpi_pct", "enroll_black_pct", 
          "enroll_white_pct", "enroll_twomore_pct","enroll_el_pct", "enroll_elserv_pct", "enroll_sped_pct",
          "dist_percent_white", "dist_percent_latinx", "dist_percent_black",
          "dist_percent_twomore", "dist_percent_asianam", "dist_percent_nhpi",
          "dist_percent_aian", "dist_percent_el", "dist_percent_sped"))
```