---
title: "Harassment and Bullying"
author: "Havisha Khurana"
date: "10/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Exploring Harassment and Bullying Data set

```{r load packages, include = FALSE}
library(rio)
library(readxl)
library(tidyverse)
library(here)
library(skimr)
```

Let's explore the data set.

```{r}
hnb <- import(here("data","Harassment and Bullying.csv")) %>% 
    as_tibble()
head(hnb)
```
There are 145 columns; removing the details of school (8 variables), there remaining `r 145-8` are combinations of intent, ethnicity, gender, ability status and English learner status.

Let's explore the unique values in different columns.

```{r unique column values}
hnb_wi <- hnb %>% 
    select(starts_with("SCH_"))

#apply(hnb_wi, 2, function(x) (unique(x)))

```
```{r unique}
apply(hnb_wi[2:126], 2, function(x) unique(x))
```

Question to the group: there are 4 negative numbers in the data: -11, -9, -6, -5. What could it mean? A substitute for data unavailable, or a code for some issues in getting that data. Anyhow, it looks like missing data and I'll start by cleaning the data set such that a column with any negative number will be removed.

**Figure this out later:**
{r remove negative columns}
hnb_clean <- hnb
apply(hnb_clean[9:145], 2, function(x)  {ifelse(x < 0, NA, x)})
hnb_clean <- hnb_clean[complete.cases(hnb_clean),]
dim(hnb_clean)

Let's do some plots:
```{r plot}
ggplot(hnb_wi) +
    geom_point(mapping = aes(x = SCH_HBALLEGATIONS_SEX, y = SCH_HBALLEGATIONS_RAC)) 
```


```{r summary}
skim(hnb_wi)

```
 
 