---
output: html_document
---
<style type="text/css">
  body{
  font-size: 12pt;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<center> <h3> EDLD 651: Introduction to Educational Data Science </h3> </center>
<center> <h3> Final Project: Outline </h3> </center>
<center> <h3> Group: Cassie, Errol, Havi, Manuel, Merly </h3> </center>

<br>
<br>

### 1. _Description of the Data_

This research project will be conducted using High School and Middle School level data from the [**Civil Rights Data Collection (CRDC)**](https://ocrdata.ed.gov/), 2017-2018 data set. The CRDC contains multiple data sets that cover a range of topics. From the overall data set, we are focusing on frequency data reported in the following areas: 

1. School Support (count of enrolled novice teachers, sworn LEO)
2. Harassment and Bullying (count of incidents of students by subgroup)
3. Referrals & Arrests (count of referrals and arrests of students by subgroup)
4. Expulsions (count of expulsions of students by subgroup)
5. Enrollment (count of enrolled students by subgroup)
6. School characteristics (Supplementary data on reporting schools and districts)

<br>

### 2. _Discussion of Preparatory Work_


#### 2.1 Getting to know the Data

As a first step, we have divided the 6 data sets among ourselves. Each person is responsible to do a basic data cleaning: 

* Select columns
* Check for missing/malformed data
* Check for duplicated observations
* Summary statistics (skimr)
* Record observations
* Create any calculated data columns

Once we get to know the data, we will try to link each data set according to our variables of interest (Research Questions). 

```{r work distribution, echo = FALSE, result = 'asis', warning = FALSE}
library(knitr)
data <- c("School Support", "Harassment and Bullying", "Referrals & Arrests", "Expulsions", "Enrollment", "School characteristics")
main <- c("Errol", "Havi", "Cassie", "Merly", "Manuel", "Errol") 
reviewer <- c("Merly", "Cassie", "Errol", "Manuel", "Havi", "Havi")
wd <- data.frame(data_set = data, lead = main, review = reviewer)
kable(wd, caption = "Data set distribution")
```

<br>

#### 2.2 Research Questions  
**1. Relationship of Concentration of novice teachers and rate of exclusionary discipline (i.e., suspensions, expulsions) in high school students** 

* Vary by student subgroups? (i.e. race, disability)
* Vary district?
* Vary by state?

**2. Relationship of security guard density and harassment and bullying or referrals + arrests (or restraint & seclusion)?** 

* Should we see if there is any correlation between corporal punishment and security guard presence?
* Might vary by student characteristic (e.g., students with disabilities, gender, grade level, or race)
* Number of transfers out of schools with guards might be at an increased level

**(_Still considering_) Discipline concentration by school level?**

For now, we can use all data regardless of level, and we will determine whether this is feasible in following meetings.


#### 2.3 Role allocation of each member


* Coordinator : Manuel 
* Project Manager 1: Merly (#1 Research question : Relationship of Concentration of novice teachers and rate of exclusionary discipline for Middle and High School Students)
* Project Manager 2. Cassie (#2 Research question: Relationship of security guard density and harassment and bullying or referrals + arrests?
* Scrum Master: Errol 
* Quality assurance (course deliverable): Havi

<br>

### Section 3: _Requirements of the Project_

* For each research question, we'll include an abstract, introduction, methods, results, discussion and references. 
* We are using multiple data sets and specific variables in this study. As a part of tidying our data set, we'll use all the required functions, i.e., `pivot_longer()`, `mutate()`, `select()`, `filter()`, `pivot_wider()`, `group_by()`, `summarize()`
* Each research question will be accompanied with figures and discussion to determine whether a relationship exists
* APA style tables, graphs and references will be included as needed.

