---
title: "Expulsion"
author: "Merly Klaas"
date: "10/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rio)
library(ggplot2)
library(readr)
library(here)
library(janitor)
library(skimr)
#library(dplyr)
#library(purrr)
```

Importing the dataset

```{r}
expl <- read.csv(here("data", "Expulsions.csv")) %>% 
  clean_names() %>%  
  as_tibble()
expl[expl < 0] <- NA


```

Count colSums for all columns

```{r}
expl_colsums <- expl %>% 
  select(9:142) %>% 
  colSums(., na.rm=TRUE) 
expl_colsums 
```


Dataset with no negative values across all variables. 

```{r}
expl1 <- expl
#expl1[expl1 < 0] <- NA

expl1 <- expl1 %>% 
  drop_na()
```

## Explusion of preschool children, by race/ethnicity and gender, disability status


```{r}
preschool_expl <- expl %>% 
  select(1:24)
#preschool_expl[preschool_expl < 0] <- NA

preschool_expl <- preschool_expl %>% 
  drop_na()
```

## Expulsion of students with educational services, by race/ethnicity, gender, disability status

```{r}
w_edu_expl <- expl %>% 
  select (1:8, 29:48, 85:103)
#w_edu_expl[w_edu_expl <0] <- NA

w_edu_expl <- w_edu_expl %>% 
  drop_na()
```

## Expulsion of students without educational services, by race/etchnicity, gender, disability status

```{r}
wo_edu_expl <- expl %>% 
  select (1:8, 49:64, 104:122)
#wo_edu_expl[wo_edu_expl <0] <- NA
wo_edu_expl <- wo_edu_expl %>% 
  drop_na()
```

## Expulsion of students under zero tolerance policy

```{r}
zero_pol_expl <- expl %>% 
  select(1:8, 65:84, 123:142)
#zero_pol_expl[zero_pol_expl <0] <- NA
zero_pol_expl <- zero_pol_expl %>% 
  drop_na()
```

```{r}
b <- expl
b[is.na(b)] <- 0
b <- b %>% 
  summarize (male_female = tot_psdisc_exp_m + tot_psdisc_exp_f + 
               tot_discwodis_expwe_m + tot_discwodis_expwe_f + 
               tot_discwodis_expwoe_m + tot_discwodis_expwoe_f + 
               tot_discwodis_expzt_m + tot_discwodis_expzt_f + 
               tot_discwdis_expwe_idea_m + tot_discwdis_expwe_idea_f +
               tot_discwdis_expwoe_idea_m + tot_discwdis_expwoe_idea_f+
                tot_discwdis_expzt_idea_m + tot_discwdis_expzt_idea_f ,
             hi= sch_psdisc_exp_hi_m + sch_psdisc_exp_hi_f + 
               sch_discwodis_expwe_hi_m + sch_discwodis_expwe_hi_f +
               sch_discwodis_expwoe_hi_m + sch_discwodis_expwoe_hi_f + 
               sch_discwdis_expwe_idea_hi_m +sch_discwdis_expwe_idea_hi_f + 
               sch_discwdis_expwoe_idea_hi_m + sch_discwdis_expwoe_idea_hi_f + 
               sch_discwodis_expzt_hi_m + sch_discwodis_expzt_hi_f + 
               sch_discwdis_expzt_idea_hi_m + sch_discwdis_expzt_idea_hi_f,
             bl= sch_psdisc_exp_bl_m + sch_psdisc_exp_bl_f + 
               sch_discwodis_expwe_bl_m + sch_discwodis_expwe_bl_f +
               sch_discwodis_expwoe_bl_m + sch_discwodis_expwoe_bl_f + 
               sch_discwdis_expwe_idea_bl_m +sch_discwdis_expwe_idea_bl_f + 
               sch_discwdis_expwoe_idea_bl_m + sch_discwdis_expwoe_idea_bl_f + 
               sch_discwodis_expzt_bl_m + sch_discwodis_expzt_bl_f + 
               sch_discwdis_expzt_idea_bl_m + sch_discwdis_expzt_idea_bl_f,
             am = sch_psdisc_exp_am_m + sch_psdisc_exp_am_f + 
               sch_discwodis_expwe_am_m + sch_discwodis_expwe_am_f +
               sch_discwodis_expwoe_am_m + sch_discwodis_expwoe_am_f + 
               sch_discwdis_expwe_idea_am_m +sch_discwdis_expwe_idea_am_f + 
               sch_discwdis_expwoe_idea_am_m + sch_discwdis_expwoe_idea_am_f + 
               sch_discwodis_expzt_am_m + sch_discwodis_expzt_am_f + 
               sch_discwdis_expzt_idea_am_m + sch_discwdis_expzt_idea_am_f,
             as= sch_psdisc_exp_as_m + sch_psdisc_exp_as_f + 
               sch_discwodis_expwe_as_m + sch_discwodis_expwe_as_f +
               sch_discwodis_expwoe_as_m + sch_discwodis_expwoe_as_f + 
               sch_discwdis_expwe_idea_as_m +sch_discwdis_expwe_idea_as_f + 
               sch_discwdis_expwoe_idea_as_m + sch_discwdis_expwoe_idea_as_f + 
               sch_discwodis_expzt_as_m + sch_discwodis_expzt_as_f + 
               sch_discwdis_expzt_idea_as_m + sch_discwdis_expzt_idea_as_f,
             hp= sch_psdisc_exp_hp_m + sch_psdisc_exp_hp_f + 
               sch_discwodis_expwe_hp_m + sch_discwodis_expwe_hp_f +
               sch_discwodis_expwoe_hp_m + sch_discwodis_expwoe_hp_f + 
               sch_discwdis_expwe_idea_hp_m +sch_discwdis_expwe_idea_hp_f + 
               sch_discwdis_expwoe_idea_hp_m + sch_discwdis_expwoe_idea_hp_f + 
               sch_discwodis_expzt_hp_m + sch_discwodis_expzt_hp_f + 
               sch_discwdis_expzt_idea_hp_m + sch_discwdis_expzt_idea_hp_f,
             wh= sch_psdisc_exp_wh_m + sch_psdisc_exp_wh_f + 
               sch_discwodis_expwe_wh_m + sch_discwodis_expwe_wh_f +
               sch_discwodis_expwoe_wh_m + sch_discwodis_expwoe_wh_f + 
               sch_discwdis_expwe_idea_wh_m +sch_discwdis_expwe_idea_wh_f + 
               sch_discwdis_expwoe_idea_wh_m + sch_discwdis_expwoe_idea_wh_f + 
               sch_discwodis_expzt_wh_m + sch_discwodis_expzt_wh_f + 
               sch_discwdis_expzt_idea_wh_m + sch_discwdis_expzt_idea_wh_f,
             tr= sch_psdisc_exp_tr_m + sch_psdisc_exp_tr_f + 
               sch_discwodis_expwe_tr_m + sch_discwodis_expwe_tr_f +
               sch_discwodis_expwoe_tr_m + sch_discwodis_expwoe_tr_f + 
               sch_discwdis_expwe_idea_tr_m +sch_discwdis_expwe_idea_tr_f + 
               sch_discwdis_expwoe_idea_tr_m + sch_discwdis_expwoe_idea_tr_f + 
               sch_discwodis_expzt_tr_m + sch_discwodis_expzt_tr_f + 
               sch_discwdis_expzt_idea_tr_m + sch_discwdis_expzt_idea_tr_f,
             lep= sch_psdisc_exp_lep_m + sch_psdisc_exp_lep_m + 
               sch_discwodis_expwe_lep_m + sch_discwodis_expwe_lep_f + 
               sch_discwodis_expwoe_lep_m + sch_discwodis_expwoe_lep_m + 
               sch_discwdis_expwe_lep_m + sch_discwdis_expwe_lep_m +
               sch_discwodis_expwoe_lep_m + sch_discwodis_expwoe_lep_f + 
               sch_discwodis_expzt_lep_m + sch_discwodis_expzt_lep_m +
               sch_discwdis_expzt_lep_m + sch_discwdis_expzt_lep_f,
             idea= sch_psdisc_exp_idea_m + sch_psdisc_exp_idea_m +
               tot_discwdis_expwe_idea_m + tot_discwdis_expwe_idea_f +
               tot_discwdis_expwoe_idea_m + tot_discwdis_expwoe_idea_f +
               tot_discwdis_expzt_idea_m + tot_discwdis_expzt_idea_m +
               sch_discwdis_expwe_504_m + sch_discwdis_expwe_504_f)
```


```{r}
b %>% 
colSums(., na.rm= TRUE)
```

Column summary for all expulsions 
```{r recoded}
cat_sum <- function(x, y) rowSums(select(x, contains(y)), na.rm = TRUE)

try <- expl %>% 
    summarise(male_female = cat_sum(.,"tot"),
              hi = cat_sum(., "_hi_"),
              bl = cat_sum(.,"_bl_"),
              am = cat_sum(.,"_am_"),
              as = cat_sum(.,"_as_"),
              hp = cat_sum (.,"_hp_"),
              wh = cat_sum(.,"_wh_"),
              tr = cat_sum(.,"_tr_"),
              lep = cat_sum(.,"_lep_"),
              idea = cat_sum(.,"_idea_")
              ) %>% 
    colSums(., na.rm= TRUE)
try
```