---
title: "Suspension"
author: "Merly Klaas"
date: "11/2/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(rio)
library(ggplot2)
library(readr)
library(here)
library(janitor)
library(skimr)
library(dplyr)
```

## Importing dataset

```{r}
susp <- read.csv(here("data", "Suspensions.csv")) %>% 
  clean_names() %>%  
  as_tibble()

```

Data set with no negative values across all variables. 

```{r}
susp1 <- susp
susp1[susp1 < 0] <- NA

susp1 <- susp1 %>% 
  drop_na()
```

## Preschool students suspension by race/ethnicity, gender, disability status
* Only one out-of-school suspension
```{r include=FALSE}
preschool_one_susp <- susp %>% 
  select (1:8, 9:28, 49:50 )
preschool_one_susp[preschool_one_susp < 0] <- NA

preschool_one_susp <- preschool_one_susp %>% 
  drop_na()
preschool_one_susp
```
* more than one out-of-school suspensions
```{r include=FALSE}
preschool_more_susp <- 
  preschool_more_susp <- susp %>% 
  select (1:8, 29:50 )
preschool_more_susp[preschool_more_susp < 0] <- NA

preschool_more_susp <- preschool_more_susp %>% 
  drop_na()
preschool_more_susp
```
Students who received one or more in-school suspension, by race/etchnicity, gender, disability status
```{r include=FALSE}
one_in_susp <- susp %>% 
  select (1:8, 51:68, 105: 124 )
one_in_susp[one_in_susp < 0] <- NA

one_in_susp<- one_in_susp %>% 
  drop_na()
one_in_susp
```

Students who received one out-of-school suspension, by race/etchnicity, gender, disability status

```{r include=FALSE}
one_out_susp <- susp %>% 
  select (1:8, 69:86, 125: 144 )
one_out_susp[one_out_susp < 0] <- NA

one_out_susp<- one_out_susp %>% 
  drop_na()
one_out_susp
```

Students who received more than one out-of-school suspension, by race/etchnicity, gender, disability status
```{r include=FALSE}

more_out_susp <- susp %>% 
  select (1:8, 87:104, 145: 167 )
more_out_susp[more_out_susp < 0] <- NA

more_out_susp<- more_out_susp%>% 
  drop_na()
more_out_susp
```
School days missed by gender, race, and disability status

```{r include=FALSE}
days_missed <- susp %>% 
  select (1:8, 168:189 )
days_missed[days_missed < 0] <- NA

days_missed <- days_missed %>% 
  drop_na()
days_missed
```


# Preschool students one suspension by Race and Gender
  ### Preschool Male Students 

```{r}
pre_one_susp_race_m <- preschool_one_susp %>% 
 summarize(bl_m = sum(sch_psdisc_singoos_bl_m)/sum(tot_psdisc_singoos_m) * 100, 
           hi_m = sum(sch_psdisc_singoos_hi_m)/sum(tot_psdisc_singoos_m)* 100,
           am_m = sum(sch_psdisc_singoos_am_m)/sum(tot_psdisc_singoos_m)* 100,
           as_m = sum(sch_psdisc_singoos_as_m)/sum(tot_psdisc_singoos_m)* 100,
           hp_m = sum(sch_psdisc_singoos_hp_m)/sum(tot_psdisc_singoos_m)* 100,
           wh_m = sum(sch_psdisc_singoos_wh_m)/sum(tot_psdisc_singoos_m)* 100,
           tr_m = sum(sch_psdisc_singoos_tr_m)/sum(tot_psdisc_singoos_m)* 100)  %>% 
  pivot_longer(cols = 
                 c(bl_m, hi_m, am_m, as_m, hp_m, wh_m,tr_m),
               names_to = "race", 
               values_to = "pre_one_susp_race_m")
pre_one_susp_race_m 

```
  
 ### Preschool one suspension Female students

```{r}
pre_one_susp_race_f  <- preschool_one_susp %>% 
 summarize(bl_f = sum(sch_psdisc_singoos_bl_f)/sum(tot_psdisc_singoos_f) * 100, 
           hi_f = sum(sch_psdisc_singoos_hi_f)/sum(tot_psdisc_singoos_f)* 100,
           am_f = sum(sch_psdisc_singoos_am_f)/sum(tot_psdisc_singoos_f)* 100,
           as_f = sum(sch_psdisc_singoos_as_f)/sum(tot_psdisc_singoos_f)* 100,
           hp_f = sum(sch_psdisc_singoos_hp_f)/sum(tot_psdisc_singoos_f)* 100,
           wh_f = sum(sch_psdisc_singoos_wh_f)/sum(tot_psdisc_singoos_f)* 100,
           tr_f = sum(sch_psdisc_singoos_tr_f)/sum(tot_psdisc_singoos_f)* 100) %>% 
  pivot_longer(cols = c(bl_f, hi_f, am_f, as_f, hp_f, wh_f,tr_f),
               names_to = "race",
               values_to = "pre_one_susp_race_f")
pre_one_susp_race_f

```

# Preschool Students with more than one suspension by race and gender

```{r}
pre_more_susp_race_m <- preschool_more_susp %>% 
  summarize(bl_m = sum(sch_psdisc_multoos_bl_m)/sum(tot_psdisc_multoos_m) * 100, 
           hi_m = sum(sch_psdisc_multoos_hi_m)/sum(tot_psdisc_multoos_m)* 100,
           am_m = sum(sch_psdisc_multoos_am_m)/sum(tot_psdisc_multoos_m)* 100,
           as_m = sum(sch_psdisc_multoos_as_m)/sum(tot_psdisc_multoos_m)* 100,
           hp_m = sum(sch_psdisc_multoos_hp_m)/sum(tot_psdisc_multoos_m)* 100,
           wh_m = sum(sch_psdisc_multoos_wh_m)/sum(tot_psdisc_multoos_m)* 100,
           tr_m = sum(sch_psdisc_multoos_tr_m)/sum(tot_psdisc_multoos_m)* 100) %>% 
  pivot_longer(cols = 
                 c(bl_m, hi_m, am_m, as_m, hp_m, wh_m,tr_m),
               names_to = "race", 
               values_to = "pre_more_susp_race_m")
pre_more_susp_race_m 

```

```{r}
pre_more_susp_race_f  <- preschool_more_susp%>% 
 summarize(bl_f = sum(sch_psdisc_multoos_bl_f)/sum(tot_psdisc_multoos_f) * 100, 
           hi_f = sum(sch_psdisc_multoos_hi_f)/sum(tot_psdisc_multoos_f)* 100,
           am_f = sum(sch_psdisc_multoos_am_f)/sum(tot_psdisc_multoos_f)* 100,
           as_f = sum(sch_psdisc_multoos_as_f)/sum(tot_psdisc_multoos_f)* 100,
           hp_f = sum(sch_psdisc_multoos_hp_f)/sum(tot_psdisc_multoos_f)* 100,
           wh_f = sum(sch_psdisc_multoos_wh_f)/sum(tot_psdisc_multoos_f)* 100,
           tr_f = sum(sch_psdisc_multoos_tr_f)/sum(tot_psdisc_multoos_f)* 100)  %>% 
  pivot_longer(cols = c(bl_f, hi_f, am_f, as_f, hp_f, wh_f,tr_f),
               names_to = "race",
               values_to = "pre_more_susp_race_f")
pre_more_susp_race_f
```


## Students without disability with one or more in-school-suspensions by race and gender

```{r include=FALSE}
in_susp_wodis_race_m <- one_in_susp %>% 
  summarize(bl_m = sum(sch_discwodis_iss_bl_m)/sum(tot_discwodis_iss_m) * 100, 
           hi_m = sum(sch_discwodis_iss_hi_m)/sum(tot_discwodis_iss_m)* 100,
           am_m = sum(sch_discwodis_iss_am_m)/sum(tot_discwodis_iss_m)* 100,
           as_m = sum(sch_discwodis_iss_as_m)/sum(tot_discwodis_iss_m)* 100,
           hp_m = sum(sch_discwodis_iss_hp_m)/sum(tot_discwodis_iss_m)* 100,
           wh_m = sum(sch_discwodis_iss_wh_m)/sum(tot_discwodis_iss_m)* 100,
           tr_m = sum(sch_discwodis_iss_tr_m)/sum(tot_discwodis_iss_m)* 100) %>% 
  pivot_longer(cols = 
                 c(bl_m, hi_m, am_m, as_m, hp_m, wh_m,tr_m),
               names_to = "race", 
               values_to = "in_susp_wodis_race_m")
in_susp_wodis_race_m 


in_susp_wodis_race_f <- one_in_susp %>%   
summarize(bl_f = sum(sch_discwodis_iss_bl_f)/sum(tot_discwodis_iss_f) * 100, 
           hi_f = sum(sch_discwodis_iss_hi_f)/sum(tot_discwodis_iss_f)* 100,
           am_f = sum(sch_discwodis_iss_am_f)/sum(tot_discwodis_iss_f)* 100,
           as_f = sum(sch_discwodis_iss_as_f)/sum(tot_discwodis_iss_f)* 100,
           hp_f = sum(sch_discwodis_iss_hp_f)/sum(tot_discwodis_iss_f)* 100,
           wh_f = sum(sch_discwodis_iss_wh_f)/sum(tot_discwodis_iss_f)* 100,
           tr_f = sum(sch_discwodis_iss_tr_f)/sum(tot_discwodis_iss_f)* 100) %>% 
  pivot_longer(cols = c(bl_f, hi_f, am_f, as_f, hp_f, wh_f,tr_f),
               names_to = "race",
               values_to = "in_susp_wodis_race_f")
  
in_susp_wodis_race_f
```

## Students with disability with one or more in-school-suspensions by race and gender
```{r}

in_susp_wdis_race_m <- one_in_susp %>%   
  summarize(bl_m = sum(sch_discwdis_iss_idea_bl_m)/sum(tot_discwdis_iss_idea_m) * 100, 
           hi_m = sum(sch_discwdis_iss_idea_hi_m)/sum(tot_discwdis_iss_idea_m)* 100,
           am_m = sum(sch_discwdis_iss_idea_am_m)/sum(tot_discwdis_iss_idea_m)* 100,
           as_m = sum(sch_discwdis_iss_idea_as_m)/sum(tot_discwdis_iss_idea_m)* 100,
           hp_m = sum(sch_discwdis_iss_idea_hp_m)/sum(tot_discwdis_iss_idea_m)* 100,
           wh_m = sum(sch_discwdis_iss_idea_wh_m)/sum(tot_discwdis_iss_idea_m)* 100,
           tr_m = sum(sch_discwdis_iss_idea_tr_m)/sum(tot_discwdis_iss_idea_m)* 100)%>% 
  pivot_longer(cols = 
                 c(bl_m, hi_m, am_m, as_m, hp_m, wh_m,tr_m),
               names_to = "race", 
               values_to = "in_susp_wdis_race_m")
in_susp_wdis_race_m

in_susp_wdis_race_f <- one_in_susp %>%   
  summarize(bl_f = sum(sch_discwdis_iss_idea_bl_f)/sum(tot_discwdis_iss_idea_f) * 100, 
           hi_f = sum(sch_discwdis_iss_idea_hi_f)/sum(tot_discwdis_iss_idea_f)* 100,
           am_f = sum(sch_discwdis_iss_idea_am_f)/sum(tot_discwdis_iss_idea_f)* 100,
           as_f = sum(sch_discwdis_iss_idea_as_f)/sum(tot_discwdis_iss_idea_f)* 100,
           hp_f = sum(sch_discwdis_iss_idea_hp_f)/sum(tot_discwdis_iss_idea_f)* 100,
           wh_f = sum(sch_discwdis_iss_idea_wh_f)/sum(tot_discwdis_iss_idea_f)* 100,
           tr_f = sum(sch_discwdis_iss_idea_tr_f)/sum(tot_discwdis_iss_idea_f)* 100) %>% 
  pivot_longer(cols = c(bl_f, hi_f, am_f, as_f, hp_f, wh_f,tr_f),
               names_to = "race",
               values_to = "in_susp_wdis_race_f")
in_susp_wdis_race_f

```

# Students without disability with only one out-of-school-suspensions by race and gender

```{r}

one_out_susp_wodis_race_m <- one_out_susp %>% 
  summarize(bl_m = sum(sch_discwodis_singoos_bl_m)/sum(tot_discwodis_singoos_m) * 100, 
           hi_m = sum(sch_discwodis_singoos_hi_m)/sum(tot_discwodis_singoos_m)* 100,
           am_m = sum(sch_discwodis_singoos_am_m)/sum(tot_discwodis_singoos_m)* 100,
           as_m = sum(sch_discwodis_singoos_as_m)/sum(tot_discwodis_singoos_m)* 100,
           hp_m = sum(sch_discwodis_singoos_hp_m)/sum(tot_discwodis_singoos_m)* 100,
           wh_m = sum(sch_discwodis_singoos_wh_m)/sum(tot_discwodis_singoos_m)* 100,
           tr_m = sum(sch_discwodis_singoos_tr_m)/sum(tot_discwodis_singoos_m)* 100) %>% 
  pivot_longer(cols = 
                 c(bl_m, hi_m, am_m, as_m, hp_m, wh_m,tr_m),
               names_to = "race", 
               values_to = "one_out_susp_wodis_race_m")
one_out_susp_wodis_race_m 

one_out_susp_wodis_race_f <- one_out_susp %>% 
  summarize(bl_f = sum(sch_discwodis_singoos_bl_f)/sum(tot_discwodis_singoos_f) * 100, 
           hi_f = sum(sch_discwodis_singoos_hi_f)/sum(tot_discwodis_singoos_f)* 100,
           am_f = sum(sch_discwodis_singoos_am_f)/sum(tot_discwodis_singoos_f)* 100,
           as_f = sum(sch_discwodis_singoos_as_f)/sum(tot_discwodis_singoos_f)* 100,
           hp_f = sum(sch_discwodis_singoos_hp_f)/sum(tot_discwodis_singoos_f)* 100,
           wh_f = sum(sch_discwodis_singoos_wh_f)/sum(tot_discwodis_singoos_f)* 100,
           tr_f = sum(sch_discwodis_singoos_tr_f)/sum(tot_discwodis_singoos_f)* 100) %>% 
  pivot_longer(cols = c(bl_f, hi_f, am_f, as_f, hp_f, wh_f,tr_f),
               names_to = "race",
               values_to = "one_out_susp_wodis_race_f")
one_out_susp_wodis_race_f

```

# Students with disability with one out-of-school-suspensions by race and gender

```{r}

one_out_susp_wdis_race_m <- one_out_susp %>% 
  summarize(bl_m = sum(sch_discwdis_singoos_idea_bl_m)/sum(tot_discwdis_singoos_idea_m) * 100, 
           hi_m = sum(sch_discwdis_singoos_idea_hi_m)/sum(tot_discwdis_singoos_idea_m)* 100,
           am_m = sum(sch_discwdis_singoos_idea_am_m)/sum(tot_discwdis_singoos_idea_m)* 100,
           as_m = sum(sch_discwdis_singoos_idea_as_m)/sum(tot_discwdis_singoos_idea_m)* 100,
           hp_m = sum(sch_discwdis_singoos_idea_hp_m)/sum(tot_discwdis_singoos_idea_m)* 100,
           wh_m = sum(sch_discwdis_singoos_idea_wh_m)/sum(tot_discwdis_singoos_idea_m)* 100,
           tr_m = sum(sch_discwdis_singoos_idea_tr_m)/sum(tot_discwdis_singoos_idea_m)* 100) %>% 
  pivot_longer(cols = 
                 c(bl_m, hi_m, am_m, as_m, hp_m, wh_m,tr_m),
               names_to = "race", 
               values_to = "one_out_susp_wdis_race_m")
one_out_susp_wdis_race_m 

one_out_susp_wdis_race_f <- one_out_susp %>% 
  summarize(bl_f = sum(sch_discwdis_singoos_idea_bl_f)/sum(tot_discwdis_singoos_idea_f) * 100, 
           hi_f = sum(sch_discwdis_singoos_idea_hi_f)/sum(tot_discwdis_singoos_idea_f)* 100,
           am_f = sum(sch_discwdis_singoos_idea_am_f)/sum(tot_discwdis_singoos_idea_f)* 100,
           as_f = sum(sch_discwdis_singoos_idea_as_f)/sum(tot_discwdis_singoos_idea_f)* 100,
           hp_f = sum(sch_discwdis_singoos_idea_hp_f)/sum(tot_discwdis_singoos_idea_f)* 100,
           wh_f = sum(sch_discwdis_singoos_idea_wh_f)/sum(tot_discwdis_singoos_idea_f)* 100,
           tr_f = sum(sch_discwdis_singoos_idea_tr_f)/sum(tot_discwdis_singoos_idea_f)* 100) %>% 
  pivot_longer(cols = c(bl_f, hi_f, am_f, as_f, hp_f, wh_f,tr_f),
               names_to = "race",
               values_to = "one_out_susp_wdis_race_f")
one_out_susp_wdis_race_f
```

# Student without disability with more than one out-of-school-suspensions by race and gender

```{r}
more_out_susp_wodis_race_m <- more_out_susp %>% 
  summarize(bl_m = sum(sch_discwodis_multoos_bl_m)/sum(tot_discwodis_multoos_m) * 100, 
           hi_m = sum(sch_discwodis_multoos_hi_m)/sum(tot_discwodis_multoos_m)* 100,
           am_m = sum(sch_discwodis_multoos_am_m)/sum(tot_discwodis_multoos_m)* 100,
           as_m = sum(sch_discwodis_multoos_as_m)/sum(tot_discwodis_multoos_m)* 100,
           hp_m = sum(sch_discwodis_multoos_hp_m)/sum(tot_discwodis_multoos_m)* 100,
           wh_m = sum(sch_discwodis_multoos_wh_m)/sum(tot_discwodis_multoos_m)* 100,
           tr_m = sum(sch_discwodis_multoos_tr_m)/sum(tot_discwodis_multoos_m)* 100) %>% 
  pivot_longer(cols = 
                 c(bl_m, hi_m, am_m, as_m, hp_m, wh_m,tr_m),
               names_to = "race", 
               values_to = "more_out_susp_wodis_race_m")
more_out_susp_wodis_race_m 

more_out_susp_wodis_race_f <- more_out_susp %>% 
  summarize(bl_f = sum(sch_discwodis_multoos_bl_f)/sum(tot_discwodis_multoos_f) * 100, 
           hi_f = sum(sch_discwodis_multoos_hi_f)/sum(tot_discwodis_multoos_f)* 100,
           am_f = sum(sch_discwodis_multoos_am_f)/sum(tot_discwodis_multoos_f)* 100,
           as_f = sum(sch_discwodis_multoos_as_f)/sum(tot_discwodis_multoos_f)* 100,
           hp_f = sum(sch_discwodis_multoos_hp_f)/sum(tot_discwodis_multoos_f)* 100,
           wh_f = sum(sch_discwodis_multoos_wh_f)/sum(tot_discwodis_multoos_f)* 100,
           tr_f = sum(sch_discwodis_multoos_tr_f)/sum(tot_discwodis_multoos_f)* 100)  %>% 
  pivot_longer(cols = c(bl_f, hi_f, am_f, as_f, hp_f, wh_f,tr_f),
               names_to = "race",
               values_to = "more_out_susp_wodis_race_f")
more_out_susp_wodis_race_f

```

Student with disability with more than one out-of-school-suspensions by race and gender

```{r}
more_out_susp_wdis_race_m <- more_out_susp %>% 
  summarize(bl_m = sum(sch_discwdis_multoos_idea_bl_m)/sum(tot_discwdis_multoos_idea_m) * 100, 
           hi_m = sum(sch_discwdis_multoos_idea_hi_m)/sum(tot_discwdis_multoos_idea_m)* 100,
           am_m = sum(sch_discwdis_multoos_idea_am_m)/sum(tot_discwdis_multoos_idea_m)* 100,
           as_m = sum(sch_discwdis_multoos_idea_as_m)/sum(tot_discwdis_multoos_idea_m)* 100,
           hp_m = sum(sch_discwdis_multoos_idea_hp_m)/sum(tot_discwdis_multoos_idea_m)* 100,
           wh_m = sum(sch_discwdis_multoos_idea_wh_m)/sum(tot_discwdis_multoos_idea_m)* 100,
           tr_m = sum(sch_discwdis_multoos_idea_tr_m)/sum(tot_discwdis_multoos_idea_m)* 100)  %>% 
  pivot_longer(cols = 
                 c(bl_m, hi_m, am_m, as_m, hp_m, wh_m,tr_m),
               names_to = "race", 
               values_to = "more_out_susp_wdis_race_m")
more_out_susp_wdis_race_m 

more_out_susp_wdis_race_f <- more_out_susp %>% 
  summarize(bl_f = sum(sch_discwdis_multoos_idea_bl_f)/sum(tot_discwdis_multoos_idea_f) * 100, 
           hi_f = sum(sch_discwdis_multoos_idea_hi_f)/sum(tot_discwdis_multoos_idea_f)* 100,
           am_f = sum(sch_discwdis_multoos_idea_am_f)/sum(tot_discwdis_multoos_idea_f)* 100,
           as_f = sum(sch_discwdis_multoos_idea_as_f)/sum(tot_discwdis_multoos_idea_f)* 100,
           hp_f = sum(sch_discwdis_multoos_idea_hp_f)/sum(tot_discwdis_multoos_idea_f)* 100,
           wh_f = sum(sch_discwdis_multoos_idea_wh_f)/sum(tot_discwdis_multoos_idea_f)* 100,
           tr_f = sum(sch_discwdis_multoos_idea_tr_f)/sum(tot_discwdis_multoos_idea_f)* 100)%>% 
  pivot_longer(cols = c(bl_f, hi_f, am_f, as_f, hp_f, wh_f,tr_f),
               names_to = "race",
               values_to = "more_out_susp_wdis_race_f")
more_out_susp_wdis_race_f

```


## Combining all percentage dataset based on gender


```{r}
all_race_m <- left_join(pre_one_susp_race_m, pre_more_susp_race_m) %>% 
  left_join(in_susp_wdis_race_m) %>% 
  left_join(one_out_susp_wodis_race_m) %>% 
  left_join(one_out_susp_wdis_race_m) %>% 
  left_join(more_out_susp_wodis_race_m) %>% 
  left_join(more_out_susp_wdis_race_m)
  
all_race_f <- left_join(pre_one_susp_race_f,pre_more_susp_race_f) %>% 
  left_join(in_susp_wdis_race_f) %>% 
  left_join(one_out_susp_wodis_race_f) %>% 
  left_join(one_out_susp_wdis_race_f) %>% 
  left_join(more_out_susp_wodis_race_f) %>% 
  left_join(more_out_susp_wdis_race_f)
```

Plot 

```{r}
all_race_male <- cbind(all_race_m[1], stack(all_race_m[2:8])) %>% 
  select(race, ind, values) %>% 
  mutate_at(vars(values), funs(round(., 1)))

all_race_female <- cbind(all_race_f[1], stack(all_race_f[2:8])) %>% 
  select(race, ind, values) %>% 
  mutate_at(vars(values), funs(round(., 1)))

```

Plot for male

```{r Plot-race-suspensions, echo=FALSE, fig.height=10, fig.width=10}
all_race_male %>% 
  ggplot(aes(fill= race, x=ind, y=values))+
  geom_bar(position = "stack", stat= "identity", color = "black", width = 0.9)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  geom_text(aes(label = paste0(values, "%")),
            position = position_stack(vjust = 0.5), size = 2) + 
  labs(x = "Suspensions", 
       y = "Percentage",
       title = "Male Students Receiving Suspension, by Race, Gender, and Disability")

```

Plot for female
```{r echo=FALSE, fig.height=10, fig.width=10}
all_race_female %>% 
  ggplot(aes(fill= race, x=ind, y=values))+
  geom_bar(position = "stack", stat= "identity", color = "black", width = 0.9)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  geom_text(aes(label = paste0(values, "%")),
            position = position_stack(vjust = 0.5), size = 2)+ 
    labs(x = "Suspensions", 
       y = "Percentage",
       title = "Female Students Receiving Suspension, by Race, Gender, and Disability",
       color = "Race")
```
  
