---
title: "Tidy Data"
author: "Havisha Khurana"
date: "11/20/2021"
output: 
    html_document:
        toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      error = FALSE,
                      cache = TRUE)
pacman::p_load(tidyverse, rio, here, janitor)
#tidylog, , ggridges, ggthemes, skimr,, reshape2, psych
```

## Enrollment Data Set

Read file

```{r load enroll file}
enroll <- import(here("data", "Enrollment.csv")) %>%
    as_tibble() %>%
    clean_names()

```

Change negative to NA

```{r enroll remove neg}
#Function to remove negative numbers from data

remove_neg <- function (x) ifelse(x < 0, NA, x)

enroll <- enroll %>% 
    mutate(across(c(9:123), remove_neg)) %>% 
    select(-c(9:31))  #removing pre-K enrollment data
```

Now, I'm using PivotLonger and PivotWider to extract all information from columns.

```{r enroll tidy_1}
#Extracting information for columns starting with 'tot_'

enroll_total <- enroll %>% 
        select(c(1:8), matches('tot'))%>% 
        pivot_longer(
            cols = -c(1:8), 
            names_to = c("enroll_type", "gender"),
            names_sep = "_",
            names_prefix = "tot_",
            values_to = "students" 
        ) %>%
    group_by(combokey, enroll_type) %>% 
        summarize(
            total = sum(students, na.rm = TRUE)
        ) %>% 
    pivot_wider(
        names_from = 2,
        values_from = total,
        names_sep = "_",
        names_prefix = "tot_"
    ) %>% 
  as_tibble()

```


```{r enroll tidy_2}
#Extracting information for columns starting with 'sch_'

enroll_tidy <- enroll %>% 
        select(-matches('tot'))%>% 
        pivot_longer(
            cols = -c(1:8), 
            names_to = c("enroll_type", "classification","gender"),
            names_sep = "_",
            names_prefix = "sch_",
            values_to = "students" 
        ) %>%         
        group_by(combokey, enroll_type, classification) %>% 
        summarize(
            total = sum(students, na.rm = TRUE)
        ) %>% 
      pivot_wider(
        names_from = c(2:3),
        values_from = total,
        names_sep = "_",
        names_prefix = "sch_"
    ) %>% 
    as_tibble()

```

```{r enroll joining}
#Joining all files to have information for analysis

enroll_final <- enroll_tidy %>% 
    right_join(.,enroll[,1:8],by = 'combokey') %>% 
    right_join(., enroll_total,by = 'combokey') %>%
    select(c(lea_state:jj), combokey, 'tot_enr',contains("_enr_"), everything()) %>% 
    as_tibble()
```


```{r enrol pct}
# Expressing enrollment across categories as percentage of total enrollment

enroll_final <- enroll_final %>%
  mutate(across(-c(1:9), ~ . / !! enroll_final$tot_enr * 100)) 
```

Note: For 11 schools, total enrollment is 0. Therefore, the percentages come as NA for them. Probably because these schools only have preK enrollment. 

```{r enroll errors}
enroll_100 <- enroll_final %>% 
  select(c(1:19)) %>% 
  filter_at(vars(c(10:19)),any_vars(.>100))
```

Note: For `r nrow(enroll_100)`, the pct over one sub-group is greater than 100 - this mostly happens in the `idea` enrollment. 

## Suspension Data set

```{r load suspension data and susset}
susp <- import(here("data", "Suspensions.csv")) %>%
    as_tibble() %>%
    clean_names() %>% 
    mutate(across(c(9:189), remove_neg))

#preK suspension data
susp_ps <- susp %>% 
  select(c(1:50)) 

#K12 suspension
susp_k12 <- susp %>% 
  select(c(1:8,51:167))

```

Tidy data for regression

```{r susp_sum}

#K-12 summary data
susp_sum <- susp_k12 %>% 
  select(combokey, matches('tot|_504_')) %>% 
  mutate(
    tot_susp = rowSums(across(where(is.numeric)), na.rm = TRUE)
  ) %>% 
  select(combokey, tot_susp)
  

```

##School Support Import and Tidying

```{r load_support_data}
support<- read_csv(here("data", "School Support.csv")) %>%
  as_tibble() %>%
  clean_names() %>% 
  mutate(across(c(9:22),remove_neg)) %>% 
  mutate(sum_new_teachers = (sch_fteteach_sy + sch_fteteach_fy),
    pct_new_teachers = sum_new_teachers/sch_teachers_curr_tot,
        )

support_tidy <- support %>% 
  pivot_longer(cols = -c(1:8,23,24),
               names_to = c("untidy1","detail1","detail2"),
               values_to = "count_teachers",
               names_prefix = "sch_",
               names_sep = "_"
  ) %>% 
  mutate(
    fte = case_when(
      str_detect(untidy1,"fte") ~ "Yes",
      TRUE ~ "No"
    ),
    untidy1 = str_remove(untidy1,"fte"),
    tot = case_when(
      str_detect(detail1,"tot") ~ "Yes",
      str_detect(detail2,"tot") ~ "Yes",
      TRUE ~ "No"
    ),
    detail1 = str_remove(detail1,"tot"),
    employee_category = untidy1,
  ) %>% 
  filter(tot == "Yes" & fte == "Yes") %>% 
  select(-c(detail2,detail1,untidy1,tot,fte,employee_category)) %>% 
  select(-c(1:6,8))


          
```


## School Characteristics Tidying
```{r load and tidy characteristics}

sch_characteristics_raw <- import(here("data","School Characteristics.csv")) %>% 
  as_tibble() %>% 
  clean_names()
sch_characteristics <- sch_characteristics_raw %>%
  mutate(across(c(9:32),remove_neg)) %>% 
  select(-c(9:26,31,32))

characteristics_tidy <- sch_characteristics %>% 
  pivot_longer(cols = (8:12),
               names_to = "type",
               values_to = "status",
               names_prefix = "sch_status_") %>% 
  pivot_wider(names_from = type,
              values_from = status) %>% 
  select(-(1:6))
```

```{r susp tidy, eval = FALSE}
# Long data for descriptive analysis
#incomplete

susp_ps_tidy <- susp_ps %>% 
        select(-matches('tot'))%>% 
        pivot_longer(
            cols = -c(1:8, 45,46), 
            names_to = c("type", "classification","gender"),
            names_sep = "_",
            names_prefix = "sch_psdisc_",
            values_to = "students" 
        )

susp_k12_tidy <- susp_k12 %>% 
        select(-matches('tot'))%>% 
        pivot_longer(
            cols = -c(1:8), 
            names_to = c("disability", "type","classification", "gender"),
            names_sep = "_",
            names_prefix = "sch_disc_",
            values_to = "students" 
        )
```

## Expulsion Data Set

```{r load expulsion data and subset}
expl <- import(here("data", "Expulsions.csv")) %>%
    as_tibble() %>%
    clean_names() %>% 
    mutate(across(c(9:142), remove_neg))

#preK expulsion data
expl_ps <- expl %>% 
  select(c(1:8), matches('psdisc')) 

#K12 data expulsion data
expl_k12 <- expl %>% 
  select(c(1:142), -matches('psdisc'))

```

Tidy data for regression

```{r expl_sum }

#K-12 summary data
expl_sum <- expl_k12  %>% 
  select(combokey, matches('tot|_504_')) %>% 
  mutate(
    tot_expl = rowSums(across(where(is.numeric)), na.rm = TRUE)
  ) %>% 
  select(combokey, tot_expl)
  
```

## Referals and Arrest Data Set

```{r ra_sum}
ra_sum <- import(here("data", "Referrals and Arrests.csv")) %>%
    as_tibble() %>%
    clean_names() %>% 
    mutate(across(c(9:84), remove_neg)) %>% 
    select(combokey, matches('tot|_504_')) %>% 
    mutate(
    tot_ref = rowSums(select(., contains('_ref_')), na.rm = TRUE),
    tot_arr = rowSums(select(., contains('_arr_')), na.rm = TRUE)
    ) %>% 
  select(combokey, tot_ref, tot_arr)
```

## Joined Data

```{r joined}
# Function to calculate rates per 1000 student
rate <- function(x) (x*1000 / enroll_final$tot_enr)

data_join <- enroll_final %>% 
  select(c(1:9, 11:15, 18:19, 17, 16, 10)) %>% 
#  left_join(susp_sum, by = 'combokey') %>% 
  left_join(expl_sum, by = 'combokey') %>% 
  left_join(ra_sum, by = 'combokey') %>% 
  left_join(support_tidy, by = "combokey") %>% 
  left_join(characteristics_tidy, by = "combokey") %>% 
  mutate(
#  susp_std = rate(tot_susp),
  expl_std = rate(tot_expl),
  ref_std = rate(tot_ref),
  arr_std = rate(tot_arr),
  new_teacher_std = rate(sum_new_teachers)
  )



```


```{r export}
#export(data_join)
```

